Rusty Linux: Advances in Rust for Linux Kernel Development

### 2024-08-30 paper.tex
Update on Overleaf.

\documentclass[sigconf]{acmart}
%\documentclass[sigconf,review,anonymous]{acmart}
\usepackage{listings, listings-rust}
\usepackage{xcolor}
\usepackage[raggedrightboxes]{ragged2e}

\pagestyle{plain}

\NewDocumentCommand{\codeword}{v}{%
\texttt{\textcolor{blue}{#1}}
}

\lstset{language=Rust, keywordstyle={\bfseries \color{blue}}}

%% Rights management information.
\setcopyright{acmlicensed}
\copyrightyear{2024}
\acmYear{2024}
\acmDOI{XXXXXXX.XXXXXXX}
%\acmSubmissionID{123-A56-BU3}

\begin{document}
%\title{Kernel Development Using Rust: A Systematic Literature Review}
\title{Rusty Linux: Advances in Rust for Linux Kernel Development}

%\author{Anonymous Authors}

@@

%\renewcommand{\shortauthors}{Panter et al.}

@@

  \textit{\textbf{Context:}} The integration of Rust into kernel development is a transformative endeavor aimed
  at enhancing system security and reliability by leveraging Rust's strong memory safety
  guarantees. 
  %This systematic literature review examines the current state of research and practice regarding the use of Rust in the kernel development.
  \textit{\textbf{Objective:}} We aim to find the current advances in using Rust in Kernel
  development to reduce the number of memory safety
  vulnerabilities in one of the most critical pieces of software that underpins all modern
  applications.
  \textit{\textbf{Method:}} By analyzing a broad spectrum of studies, we identify the advantages Rust offers,
  %such as reducing vulnerabilities like buffer overflows and use-after-free errors, which are
  %prevalent in C-based kernel development. The review also 
  highlight the challenges faced,
  %including integration hurdles with existing C codebases, performance considerations, 
  and emphasise the need for community consensus on Rust's adoption.
  \textit{\textbf{Results:}} Our findings suggest that while the initial implementations of Rust in the kernel
  show promising results in terms of safety and stability, significant challenges remain. These challenges
  include achieving seamless interoperability with existing kernel components, maintaining
  performance, and ensuring adequate support and tooling for developers.
  \textit{\textbf{Conclusions:}} This study 
  %provides a comprehensive overview of the
  %potential and challenges of incorporating Rust into kernel development. It 
  underscores the need for
  continued research and practical implementation efforts to fully realize the benefits of Rust. By
  addressing these challenges, the integration of Rust could mark a significant step forward in the
  evolution of operating system development towards safer and more reliable systems.

@@

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002944.10011122.10002945</concept_id>
<concept_desc>General and reference~Surveys and overviews</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011006.10011008.10011024</concept_id>
<concept_desc>Software and its engineering~Language features</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10010520.10010521</concept_id>
<concept_desc>Computer systems organization~Architectures</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{General and reference~Surveys and overviews}
\ccsdesc[500]{Software and its engineering~Language features}
\ccsdesc[500]{Computer systems organization~Architectures}

@@

% \received{30 May 2024}
% \received[revised]{TBD}
\received[accepted]{22 July 2024}

@@

 \begin{figure*}
    \centering
     \includegraphics[width=7in]{process-diagram.pdf}
     \caption{Process Diagram for Paper selection}
     \label{fig:process}
 \end{figure*}

@@

\begin{table}
\begin{tabular}{|| p{2cm}| p{4cm} | p{1cm} ||}
 \hline
 Database & Query & Results \\
 \hline\hline
 ACM  & (Abstract:("operating system" kernel linux OS) AND Abstract:(rust) )OR (Title:("operating
 system" OS kernel linux) AND Title:(rust))  & 30 \\
 IEEE & ("All Metadata":"operating system" AND "All Metadata":rust) OR ("All Metadata":kernel AND
 "All Metadata":rust) OR ("All Metadata":linux AND "All Metadata":rust) & 39 \\
 Scopus & TITLE-ABS-KEY ( ( "operating system" OR kernel OR linux ) AND rust ) AND PUBYEAR > 2018
 AND PUBYEAR < 2025 AND ( LIMIT-TO ( SUBJAREA , "COMP" ) ) AND ( LIMIT-TO ( LANGUAGE , "English" ) )
 & 87 \\
 Springer Link & Operating System Rust Kernel (Conference Paper, Article, Research article) Subdiscipline: Software engineering/programming and operating systems & 29 \\
 \hline
\end{tabular}
\caption{Search Queries used for each Database.}
\label{tab:keywords}
\end{table}

@@

\begin{table*}
    \begin{tabular}{||l|l|l||}
    \hline
    Approach & Papers & Operating System in Rust\\
    \hline\hline
    Monolithic  & \cite{The_kernel_development_community_undated-iw}, \cite{Li2019-ru}, \cite{Miller2021-pg}, \cite{Oikawa2023-ms} & \href{https://docs.kernel.org/rust/}{Linux kernel v6.1+}\\
    Micro-kernel & \cite{Chen2023-wb},\cite{Liang2021-bo}, \cite{Liu2024-xe}, \cite{Narayanan2020-gs}, \cite{Narayanan2019-fd} & Atmosphere, Redox, Redleaf\\
    Embedded & \cite{Culic2022-bk}, \cite{Vishnunaryan2022-yd} & \href{https://github.com/tock/tock}{Tock}, \href{https://hubris.oxide.computer/}{Hubris}, \href{https://www.drone-os.com/}{Drone}, \href{https://bern-rtos.org/}{Bern}, HarSaRK \\
    Unikernel & \cite{Lankes2019-cm},  \cite{Boos2020-zh}, \cite{Ijaz2023-da}, \cite{Sung2020-bb}  & RustyHermit, Theseus \\
    Exokernel & \cite{Li2024-yb} & W-Kernel \\
    \hline
  \end{tabular}
  \caption{Approaches and Methodologies for Rust in the Kernel}
    \label{tab:RQ1}
\end{table*}

@@

\begin{table}
    \begin{tabular}{||l|l|l||}
    \hline
    No. & Implication & Studies that Reported the challenge\\
    \hline\hline
    1 & Performance & \cite{Gonzalez2023-ek}, \cite{Li2024-be}, \cite{Ma2023-ef}\\
    2 & Throughput & \cite{Gonzalez2023-ek}\\
    3 & Latency & \cite{Culic2022-bk} \\
    \hline
  \end{tabular}
  \caption{Performance Implications of Rust in the Kernel}
    \label{tab:RQ2}
\end{table}

@@

\begin{table*}
    \begin{tabular}{p{1cm} p{5cm} p{10cm}}
        \hline
        No. & Challenge & Description\\
        \hline
        1 & Binary Size~\cite{Ayers2022-sf}  &
        \begin{itemize}
            \item Deeply ingrained monomorphization which increased the size of the rust binaries
            \item Compiler optimizations are not as mature as some C based compilers, thus increasing binary size
            \item Hidden data structures and data
            \item Sub-optimal compiler generated support code
        \end{itemize}
        \\
        \hline
        2 & Missing Features~\cite{Burtsev2021-mh} &
        \begin{itemize}
            \item Supporting trait bounds on functions and closures with any number of arguments
            \item Expose type information in procedural macros
            \item Support a collision-free, unique type identifier

            \item Support typed assembly language for Rust
            \item Support trusted build environments
            \item Provide software-only stack guard with extensible probing interface
            \item Develop zero-copy serialization of “plain-old” data structures
        \end{itemize}
        \\
        \hline
        3 & Soundness~\cite{Klimt2023-ob} &
        \begin{itemize}
            \item Unsynchronized Global State - any use of mutable statics is unsafe
            \item C-Style Abstractions - Use rust style abstractions to properly encapsulate internal unsafe usage of raw pointers
            \item Aliasing Mutable References - Giving out raw pointers to memory that is also referenced mutably
            \item Re-implementing Memory Access - instead of accessing specific memory regions in assembly, creating and using references to the whole region is preferable
        \end{itemize}
        \\
        \hline
        4 & Panics \cite{Ma2023-ef},\cite{Burtsev2021-mh} &
        \begin{itemize}
            \item Support extendable, no\_std unwind library
            \item Stack unwinding in embedded environments
        \end{itemize}
        \\
        \hline
        5 & C Interop \cite{Miller2019-xm}, \cite{Li2021-xo} &
        \begin{itemize}
            \item Kernel interfaces, while designed for extensibility, are not designed for type safety
            \item Hybrid Code Flow. The Rust compiler can not track ownership when switching between modules written in C and Rust
        \end{itemize}
        \\
    \hline
  \end{tabular}
  \caption{Challenges Unique to the Rust Programming Language}
    \label{tab:RQ3}
\end{table*}

@@

\balance

\bibliographystyle{ACM-Reference-Format}
\bibliography{slr-paper-memory-safety-rust.bib}

>>>>

\documentclass[sigconf]{acmart}
% Camera Ready documentation
% https://www.acm.org/binaries/content/assets/publications/taps/latex-best_practices-06-may-2020.pdf

%% Rights management information.
\copyrightyear{2024}
\acmYear{2024}
\setcopyright{rightsretained}
\acmConference[ESEM '24]{Proceedings of the 18th ACM / IEEE International Symposium on Empirical Software Engineering and Measurement}{October 24--25, 2024}{Barcelona, Spain}
\acmBooktitle{Proceedings of the 18th ACM / IEEE International Symposium on Empirical Software Engineering and Measurement (ESEM '24), October 24--25, 2024, Barcelona, Spain}
\acmDOI{10.1145/3674805.3690756}
\acmISBN{979-8-4007-1047-6/24/10}

\begin{document}
\title{Rusty Linux: Advances in Rust for Linux Kernel Development}

@@

@@

  \textit{\textbf{Context:}} The integration of Rust into kernel development is a transformative endeavor aimed at enhancing system security and reliability by leveraging Rust's strong memory safety guarantees. 
  \textit{\textbf{Objective:}} We aim to find the current advances in using Rust in Kerneldevelopment to reduce the number of memory safety vulnerabilities in one of the most critical pieces of software that underpins all modern applications.
  \textit{\textbf{Method:}} By analyzing a broad spectrum of studies, we identify the advantages Rust offers, highlight the challenges faced, and emphasise the need for community consensus on Rust's adoption.
  \textit{\textbf{Results:}} Our findings suggest that while the initial implementations of Rust in the kernel show promising results in terms of safety and stability, significant challenges remain. These challenges include achieving seamless interoperability with existing kernel components, maintaining performance, and ensuring adequate support and tooling for developers.
  \textit{\textbf{Conclusions:}} This study  underscores the need for
  continued research and practical implementation efforts to fully realize the benefits of Rust. By addressing these challenges, the integration of Rust could mark a significant step forward in the evolution of operating system development towards safer and more reliable systems.

@@

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10002944.10011122.10002945</concept_id>
       <concept_desc>General and reference~Surveys and overviews</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10002978.10003006.10003007</concept_id>
       <concept_desc>Security and privacy~Operating systems security</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10010940.10010941.10010949</concept_id>
       <concept_desc>Software and its engineering~Operating systems</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10011006.10011008.10011024</concept_id>
       <concept_desc>Software and its engineering~Language features</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{General and reference~Surveys and overviews}
\ccsdesc[500]{Security and privacy~Operating systems security}
\ccsdesc[300]{Software and its engineering~Operating systems}
\ccsdesc[500]{Software and its engineering~Language features}

@@

\received[accepted]{22 July 2024}

@@

 \begin{figure*}
    \centering
     \includegraphics[width=7in]{process-diagram.pdf}
     \caption{Process Diagram for Paper selection}
     \Description{The process that we followed for selecting papers}
     \label{fig:process}
 \end{figure*}

@@

\begin{table}
\begin{tabular}{|| p{2cm}| p{4cm} | p{1cm} ||}
 \hline
 Database & Query & Results \\
 \hline\hline
 ACM  & (Abstract:("operating system" OR Kernel OR linux OR OS) AND Abstract:(Rust)) OR (Title:("operating system" OR OS OR kernel OR linux) AND Title:(Rust))  & 30 \\
 IEEE & ("All Metadata":"operating system" AND "All Metadata":rust) OR ("All Metadata":kernel AND
 "All Metadata":rust) OR ("All Metadata":linux AND "All Metadata":rust) & 39 \\
 Scopus & TITLE-ABS-KEY ( ( "operating system" OR kernel OR linux ) AND rust ) AND PUBYEAR > 2018
 AND PUBYEAR < 2025 AND ( LIMIT-TO ( SUBJAREA , "COMP" ) ) AND ( LIMIT-TO ( LANGUAGE , "English" ) )
 & 87 \\
 Springer Link & Operating System Rust Kernel (Conference Paper, Article, Research article) Subdiscipline: Software engineering/programming and operating systems & 29 \\
 \hline
\end{tabular}
\caption{Search Queries used for each Database.}
\label{tab:keywords}
\end{table}

@@

\begin{table*}
    \caption{Approaches and Methodologies for Rust in the Kernel}
    \begin{tabular}{||l|l|l||}
    \hline
    Approach & Papers & Operating System in Rust\\
    \hline\hline
    Monolithic  & \cite{The_kernel_development_community_undated-iw}, \cite{Li2019-ru}, \cite{Miller2021-pg}, \cite{Oikawa2023-ms} & \href{https://docs.kernel.org/rust/}{Linux kernel v6.1+}\\
    Micro-kernel & \cite{Chen2023-wb},\cite{Liang2021-bo}, \cite{Liu2024-xe}, \cite{Narayanan2020-gs}, \cite{Narayanan2019-fd} & Atmosphere, Redox, Redleaf\\
    Embedded & \cite{Culic2022-bk}, \cite{Vishnunaryan2022-yd} & \href{https://github.com/tock/tock}{Tock}, \href{https://hubris.oxide.computer/}{Hubris}, \href{https://www.drone-os.com/}{Drone}, \href{https://bern-rtos.org/}{Bern}, HarSaRK \\
    Unikernel & \cite{Lankes2019-cm},  \cite{Boos2020-zh}, \cite{Ijaz2023-da}, \cite{Sung2020-bb}  & RustyHermit, Theseus \\
    Exokernel & \cite{Li2024-yb} & W-Kernel \\
    \hline
  \end{tabular}
  \label{tab:RQ1}
\end{table*}

@@

\begin{table}
    \caption{Performance Implications of Rust in the Kernel}
    \begin{tabular}{||l|l|l||}
    \hline
    No. & Implication & Studies that Reported the challenge\\
    \hline\hline
    1 & Performance & \cite{Gonzalez2023-ek}, \cite{Li2024-be}, \cite{Ma2023-ef}\\
    2 & Throughput & \cite{Gonzalez2023-ek}\\
    3 & Latency & \cite{Culic2022-bk} \\
    \hline
  \end{tabular}
  \label{tab:RQ2}
\end{table}

@@

\begin{table*}
    \caption{Challenges Unique to the Rust Programming Language}
    \begin{tabular}{p{1cm} p{5cm} p{10cm}}
        \hline
        No. & Challenge & Description\\
        \hline
        1 & Binary Size~\cite{Ayers2022-sf}  &
        \begin{itemize}
            \item Deeply ingrained monomorphization which increased the size of the rust binaries
            \item Compiler optimizations are not as mature as some C based compilers, thus increasing binary size
            \item Hidden data structures and data
            \item Sub-optimal compiler generated support code
        \end{itemize}
        \\
        \hline
        2 & Missing Features~\cite{Burtsev2021-mh} &
        \begin{itemize}
            \item Supporting trait bounds on functions and closures with any number of arguments
            \item Expose type information in procedural macros
            \item Support a collision-free, unique type identifier

            \item Support typed assembly language for Rust
            \item Support trusted build environments
            \item Provide software-only stack guard with extensible probing interface
            \item Develop zero-copy serialization of “plain-old” data structures
        \end{itemize}
        \\
        \hline
        3 & Soundness~\cite{Klimt2023-ob} &
        \begin{itemize}
            \item Unsynchronized Global State - any use of mutable statics is unsafe
            \item C-Style Abstractions - Use rust style abstractions to properly encapsulate internal unsafe usage of raw pointers
            \item Aliasing Mutable References - Giving out raw pointers to memory that is also referenced mutably
            \item Re-implementing Memory Access - instead of accessing specific memory regions in assembly, creating and using references to the whole region is preferable
        \end{itemize}
        \\
        \hline
        4 & Panics \cite{Ma2023-ef},\cite{Burtsev2021-mh} &
        \begin{itemize}
            \item Support extendable, no\_std unwind library
            \item Stack unwinding in embedded environments
        \end{itemize}
        \\
        \hline
        5 & C Interop \cite{Miller2019-xm}, \cite{Li2021-xo} &
        \begin{itemize}
            \item Kernel interfaces, while designed for extensibility, are not designed for type safety
            \item Hybrid Code Flow. The Rust compiler can not track ownership when switching between modules written in C and Rust
        \end{itemize}
        \\
    \hline
  \end{tabular}
  \label{tab:RQ3}
\end{table*}

@@

\bibliographystyle{ACM-Reference-Format}
\bibliography{paper.bib}
